---
title: "Eicosanoids mediators of diabetes"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
---

```{r options, echo = FALSE}
options(knitr.kable.NA = "")
knitr::opts_chunk$set(include = TRUE,
                      echo = TRUE,
                      message = FALSE,
                      results = "asis",
                      cache = FALSE,
                      warning = FALSE)
```

# Libraries

<details>
  <summary>Open/Close</summary>

```{r libraries, results = 'hide'}
library(dplyr)
library(sessioninfo)
library(knitr)
library(readr)
library(tidyr)
library(parallel)
library(ggplot2)
library(broom)
library(gridExtra)
library(gplots)
library(survival)
library(purrr)
library(survminer)
library(tibble)
library(car)
library(forestplot)
library(rlang)
library(ggpubr)
library(reshape2)
library(metafor)
```

</details>

# Session

Analyses were performed using `r sessioninfo::platform_info()$version` on 
`r sessioninfo::platform_info()$system`.

<details><summary>Packages</summary>

```{r Session info, echo = FALSE}
sessioninfo::package_info() %>%
    data.frame %>%
    filter(attached == TRUE) %>%
    select(package, loadedversion, date, source) %>%
    kable(row.names = FALSE)
```

</details>

# Source code

```{r source file list, echo = FALSE}
sourcefiles <- c("articles-importer.R",
                 "articles-utils.R",
                 "articles-models.R",
                 "articles-plots.R",
                 "articles-step.R",
                 "articles-riskscore.R")
```

```{r import source files, echo = FALSE}
for (f in sourcefiles) {
    source(f)
}
```

```{r embed files to html, echo = FALSE}
xfun::embed_files(c("eicosanoidsdm.Rmd", sourcefiles))
```

#  Folder structure

```{r clean folders}
rmdir("cache")
mkdir("cache", "report", "session", "rds")
```

# Variables

<details><summary>Included variables</summary>

```{r variables to be include, echo = FALSE}
(mvars <- tribble(~var, ~type, ~dropmissing, ~useforfiltering,
                   "Sample_ID", "categorical", TRUE, FALSE,
                   "BL_AGE", "continuous", TRUE, FALSE,
                   "female", "binomial", TRUE, FALSE,
                   "PAINO", "continuous", TRUE, FALSE,
                   "PITUUS", "continuous", TRUE, FALSE,
                   "VYOTARO", "continuous", FALSE, FALSE,
                   "LANTIO", "continuous", FALSE, FALSE,
                   "BMI", "continuous", TRUE, FALSE,
                   "WHR", "continuous", FALSE, FALSE,
                   "SYSTM", "continuous", TRUE, FALSE,
                   "DIASM", "continuous", FALSE, FALSE,
                   "HBA1C", "continuous", FALSE, FALSE,
                   "ALKI2_FR02", "continuous", FALSE, FALSE,
                   "HDL", "continuous", FALSE, FALSE,
                   "LDL_DIRECT", "continuous", FALSE, FALSE,
                   "TRIG", "continuous", TRUE, FALSE,
                   "BP_TREAT", "binomial", TRUE, FALSE,
                   "KOULGR", "binomial", FALSE, FALSE,
                   "CURR_SMOKE", "binomial", FALSE, FALSE,
                   "Q57X", "categorical", TRUE, FALSE,
                   "hsCRP", "continuous", TRUE, FALSE,
                   "DIAB_FAMILYHIST", "binomial", TRUE, FALSE,
                   "PREVAL_DIAB_T2", "binomial", TRUE, FALSE,
                   "INCIDENT_DIAB_T2", "binomial", TRUE, FALSE,
                   "DIAB_T2_AGEDIFF", "continuous", TRUE, FALSE,
                   "EAST", "binomial", TRUE, FALSE,
                   "plate", "categorical", TRUE, FALSE,
                   "pregnant", "binomial", TRUE, TRUE,
                   "PREVAL_DIAB_T1", "binomial", TRUE, TRUE,
                   "INCIDENT_DIAB_T1", "binomial", TRUE, TRUE,
                   "INCIDENT_CR_ANYCANC", "binomial", TRUE, TRUE,
                  "PREVAL_CR_ANYCANC", "binomial", TRUE, TRUE,
                  "PREVAL_CVD", "binomial", FALSE, FALSE,
                  "LIPID_TREAT", "binomial", FALSE, FALSE) %>%
      dplyr::left_join(., importinfo(), by = "var")) %>%
    kable
```

</details>

# Data import

For the analyses, exclude pregnant women, prevalent/incident type 1 diabetics
and cancers (except non-melanoma skin cancers aka ICD10 category C44).

```{r import data}
dset <- (dset.raw <- importdata()) %>%
    filter_at(vars(getmissing(mvars)), all_vars(!is.na(.))) %>%
    filter_at(vars(getfilters(mvars)), all_vars(. == 0)) %>%
    mutate_at(vars(getcategorical(mvars)), factor)
```

Dimensions before filtering are (`r dim(dset.raw)`) and after filtering (`r dim(dset)`).

<details><summary>Number of participants with missing values (NA)</summary>

```{r filtered values missing}
dset.raw %>%
    mutate_all(is.na) %>%
    count_(vars = getmissing(mvars)) %>%
    kable
```

</details>

<details><summary>Number of participants with other exclusion criteria</summary>

```{r filtered values excluded}
dset.raw %>%
    filter_at(vars(getmissing(mvars)), all_vars(!is.na(.))) %>%
    mutate_all(~. != 0) %>%
    count_(vars = getfilters(mvars)) %>%
    kable
```

</details>

# Characteristics

<details><summary>Continuous variables</summary>

```{r Characteristics continuous, echo = FALSE}
dset %>%
    dplyr::select(female, one_of(getcontinuous(mvars))) %>%
    gather(key, value, -female) %>%
    group_by(key, female) %>%
    summarize(n = n(),
              na = sum(is.na(value)),
              mean = mean(value, na.rm = TRUE),
              sd = sd(value, na.rm = TRUE),
              gmean = gmean(value, na.rm = TRUE),
              iqr = IQR(value, na.rm = TRUE)) %>%
    mutate_if(is.numeric, ~round(., 3)) %>%
    kable
```

</details>

<details><summary>Categorical variables</summary>

```{r Characteristics cagegorical, echo = FALSE}
dset %>%
    select(one_of(getcategorical(mvars))) %>%
    gather(key, value, -female) %>%
    group_by(key, female) %>%
    summarize(n = n(),
              na = sum(is.na(value)),
              zero = sum(value == 0, na.rm = TRUE),
              one = sum(value == 1, na.rm = TRUE),
              two = sum(value == 2, na.rm = TRUE),
              three = sum(value == 3, na.rm = TRUE)) %>%
    mutate_if(is.numeric, ~round(., 3)) %>%
    kable
```

</deails>


# Time-to-event analysis

Outcome: incident diabetes (not type 1) & time: time of
diagnosis. Exclude prevalent diabetes type 2. Cox-regression models
adjusted for sex, age, east/west, further with BMI, parental history
of diabetes, systolic blood pressure, triglycerides and mass
spectrometry batch; third model further with hs-CRP. 

```{r cox covariates, echo = FALSE}
( cox.covariates <- list("model_0" =  c("BL_AGE", "female", "EAST", "plate"),
                         "model_1" =  c("BL_AGE", "female", "EAST",  "plate",
                                        "BMI", "Q57X", "DIAB_FAMILYHIST", "SYSTM", "BP_TREAT", "TRIG"),
                         "model_2" =  c("BL_AGE", "female", "EAST", "plate",
                                        "BMI", "Q57X", "DIAB_FAMILYHIST", "SYSTM", "BP_TREAT", "TRIG",
                                        "PREVAL_CVD", "LIPID_TREAT"),
                         "model_3" =  c("BL_AGE", "female", "EAST", "plate",
                                        "BMI", "Q57X", "DIAB_FAMILYHIST", "SYSTM", "BP_TREAT", "TRIG",
                                        "PREVAL_CVD", "LIPID_TREAT",
                                        "hsCRP")) ) %>%
    map_df(~data.frame(covariates = paste(., collapse = ", ")), .id = "model") %>%
    kable
```

 Running Cox models

```{r cox models}
cox.dm <- lapply(cox.covariates, function(covariates)
    loop.cox(dset = dset %>% filter(PREVAL_DIAB_T2 == 0),
             response = "DIAB_T2",
             loops = getmetabolistes(dset),
             covariates = covariates))
```

<details><summary>Comparing significant associations</summary>

```{r cox significant results, echo = FALSE}
cox.dm %>%
    map_df(~loop.results(.x), .id = "model") %>%
    filter(qval < 0.05) %>%
    mutate(term = bioproperty(term),
           result = sprintf("%.3f±%.3f", estimate, std.error)) %>%
    select(model, term, result) %>%
    spread(model, result) %>%
    kable
```

</details>

<details><summary>Numerical results</summary>

```{r cox all results, echo = FALSE}
cox.dm %>%
    map_df(~loop.results(.x), .id = "model") %>%
    arrange(model, qval) %>%
    kable
```

```{r correlation model last element, echo = FALSE}
coxmodel.significant.eicosanoids <- cox.dm[["model_2"]] %>%
    loop.results %>%
    filter(qval < 0.05)
```


</details>

<details><summary>Correlation plot</summary>

```{r correlation model, echo = FALSE}
ret.correlation <- coxmodel.significant.eicosanoids$term %union% "hsCRP" %>%
    spearmancorrelation(dset = dset %>% filter(PREVAL_DIAB_T2 == 0), vars = .)
```

```{r correlation plot, echo = FALSE, results = 'hide'}
mycorplot(dset = ret.correlation,
          file = "cache/corplot.png",
          height = 1560,
          width = 1640)
```

<img src = "cache/corplot.png" />

</details>

<details><summary>Manhattan plot</summary>

```{r manhattan plot, echo = FALSE}
cox.dm[["model_2"]] %>%
    loop.results %>%
    plot.manhattanplot(., psignf = 0.05, nlabels = 8) %>%
    ggsave(file = "cache/manhattanplot.jpg",
           plot = .,
           height = 6,
           width = 7,
           dpi = 600)
```

<img src = "cache/manhattanplot.jpg" />
</details>


<details><summary>Variables failing proportionality check in last model</summary>

Checking the proportionality assumption using the Schoenfeld residuals

```{r cox check proportionality}
cox.dm[["model_2"]] %>%
    check.proportionality %>%
    map_df(~as.data.frame(.x$table) %>% tibble::rownames_to_column("term"),
           .id = "testing") %>%
    filter(p < 0.05) %>%
    kable
```

</details>

<details><summary>Deviance residuals</summary>

```{r coxmodel residuals, echo = FALSE}
ggsave(file = "cache/coxmodel-residuals.png",
       plot = grid.arrange(grobs = loop.coxresiduals(filter.significant(cox.dm[["model_2"]])),
                           ncol = 8,
                           bottom = "Index number of observer",
                           left = "Deviance residuals"),
       height = 40,
       width = 12,
       dpi = 300,
       unit = "in")
```

<img src = "cache/coxmodel-residuals.png" />

</details>


# Stepwise Cox regression

Stepwise Cox regression with forward selection: choose 5-6 eicosanoids
and build score out of them (need to check units before
combining!). Present their associations (forest plot?) and possibly
replicate in FHS and/or DILGOM.

```{r cox forward dataset}
dset.fwd <- dset %>%
    filter(PREVAL_DIAB_T2 == 0) %>%
    mutate_at(vars(getcategorical(mvars) %union% "plate" %difference% "INCIDENT_DIAB_T2"), as.numeric) %>%
    mutate_at(vars("INCIDENT_DIAB_T2"), eventasnumber)
cox.fwd.mzids <- coxmodel.significant.eicosanoids %>% pull(term)
```


Including the significant eicosanoids from the last model in the list

<details><summary>Included eicosanoids</summary>

```{r cox fowrawd significant, echo = FALSE}
coxmodel.significant.eicosanoids %>% kable
```

</details>

Running stepwise coxph using Bonferroni based p value for including/excluding new terms.

```{r cox forward, results = 'hide'}
cox.fwd <- stepwise.coxph(Time = "DIAB_T2_AGEDIFF",
                          Status = "INCIDENT_DIAB_T2",
                          variable.list = cox.fwd.mzids,
                          in.variable = cox.covariates[["model_2"]],
                          sle = 0.05/length(getmetabolistes(dset)),
                          sls = 0.05/length(getmetabolistes(dset)),
                          data = dset.fwd)
```

<details><summary>Numerical results</summary>

```{r cox forward print results, echo = FALSE}
cox.fwd %>%
    tidy(exponentiate = TRUE) %>%
    kable
```

</details>

<details><summary>Forest plot</summary>

```{r forest plot, echo = FALSE, results = 'hide'}
cox.fwd %>%
    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::filter(grepl("mzid", term)) %>%
    mutate(mean_ci = sprintf("%.2f (%.2f to %.2f)", estimate, conf.low, conf.high),
           mzrt = bioproperty(term)) %>%
    myforestplot(file = "cache/forestmap.jpg")
```

<img src = "cache/forestmap.jpg" />

</details>



# Risk score

Use quartiles for regression analyses with score, possibly replicate
in FHS and/or DILGOM. Also of interest: Kaplan-Meier curve for the
eicosanoid score using lowest and highest quartile.

Using stepwise results to define risk formula.

```{r risk formula, echo = FALSE, results = 'asis'}
getformula(cox.fwd)
```

Calculating sclaed risk score and risk classes for participant without
diabetes in baseline.

```{r risk score}
rset <- dset %>%
    filter(PREVAL_DIAB_T2 == 0) %>%
    getriskset(., cox.fwd)
```

<details><summary>Basic characteristics by risk class</summary>

```{r risk score characteristics}
rset %>%
    group_by(riskclass) %>%
    summarize(n = n(),
              age = mean(BL_AGE),
              female = sum(female == 1),
              BMI = mean(BMI),
              sys = mean(SYSTM),
              incident_dm = sum(INCIDENT_DIAB_T2 == 1),
              timed_dm = mean(DIAB_T2_AGEDIFF)) %>%
    kable
```

</details>

<details><summary>Cox models for risk score</summary>

```{r risk model riskclass, echo = FALSE}
lapply(c2l("riskclass", "riskpersd"), function(risk)
    lapply(list("basemodel" = cox.covariates[["model_2"]],
                "interaction" = cox.covariates[["model_2"]] %union%
                    paste0(risk, ":female")), function(vars)
                        loop.cox(dset = rset,
                                 response = "DIAB_T2",
                                 loops = risk,
                                 covariates = vars))) %>%
    map_df(.,
           ~map_df(.x, ~tidy(.x, exponentiate = TRUE), .id = "model"),
           .id = "risk") %>%
    kable
```

</details>

<details><summary>Kaplan-Meier</summary>

```{r kaplan mayer, echo = FALSE}
g.km <- survfit(Surv(DIAB_T2_AGEDIFF, INCIDENT_DIAB_T2) ~ riskclass, 
                data = rset %>%
                    filter(PREVAL_DIAB_T2 == 0) %>%
                    mutate_at(vars("INCIDENT_DIAB_T2"), eventasnumber)) %>%
    ggsurvplot(fit = .,
               pval = TRUE,
               pval.method = TRUE,
               pval.coord = c(12, 0.55),
               pval.method.coord = c(12, 0.60),
               legend = "right",
               legend.title = "Risk class",
               legend.labs = c("Q1", "Q2", "Q3", "Q4"),
               palette = c("black", "gray50", "gray80", "red"),
               break.time.by = 3,
               xlim = c(-0.5, 16),
               ylim = c(0.5, 1),
               xlab = "Time (years)",
               ylab = "Incident type 2 diabetes mellitus",
               risk.table = TRUE,
               risk.table.height = 0.4,
               surv.scale = "percent")

ggsave(file = "cache/km-riskclass.jpg",
       plot = print(g.km),
       height = 6,
       width = 7,
       dpi = 600)
```

<img src = "cache/km-riskclass.jpg" />

</details>

<details>
<summary>Disease-free survival</summary>

```{r disease free survival}
rset %>%
    group_by(riskclass) %>%
    summarize(event = sum(INCIDENT_DIAB_T2 == 1),
              noevent = sum(INCIDENT_DIAB_T2 == 0)) %>%
    mutate(total = event + noevent,
           dfs = 1-event/total) %>%
    kable
```

</details>


# Replication


## DILGOM replication

Importing dilgom data

```{r dilgom data import}
dset.dilgom <- (dset.dilgom.raw <- importdata(import = importdilgomdata)) %>%
    filter_at(vars(getmissing(mvars)), all_vars(!is.na(.))) %>%
    filter_at(vars(getfilters(mvars)), all_vars(. == 0)) %>%
    mutate_at(vars(getcategorical(mvars)), factor) %>%
    rename(LDL_DIRECT = LDL)
```

<details>
  <summary>Jitter plots for the three eicosanoids</summary>

```{r plot mzids dilgom, echo = FALSE}
g.scatter.dilgom <- dset.dilgom %>%
    select(starts_with("mzid_")) %>%
    gather(key, value) %>%
    ggplot(aes(x = key, y = value)) +
    geom_jitter(shape = ".") +
    ylim(-3, 4) +
    labs(title = "DILGOM") +
    theme_classic() +
    theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

```{r plot mzids fr02, echo = FALSE}
g.scatter.fr02 <- dset %>%
    select(one_of("mzid_279.196600_3.7247", "mzid_311.223100_2.9230", "mzid_335.223200_2.6455")) %>%
    gather(key, value) %>%
    ggplot(aes(x = key, y = value)) +
    geom_jitter(shape = ".") +
    ylim(-3, 4) +
    labs(title = "FR02") +
    theme_classic() +
    theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

```{r plot mzids join, echo = FALSE}
arrangeGrob(g.scatter.dilgom, g.scatter.fr02, ncol = 2) %>%
       ggsave(file = "cache/eicosanoid_scatter.png",
              plot = .,
              height = 6,
              width = 10,
              dpi = 300)
```

<img src = "cache/eicosanoid_scatter.png" />

</details>

<details>
  <summary>Box plot for the three eicosanoids</summary>

```{r dilgom eicosanoid boxplot}
( dset.dilgom %>%
    filter(PREVAL_DIAB_T2 == 0) %>%
    select(starts_with("mzid_"), INCIDENT_DIAB_T2) %>%
    gather(key, value, -INCIDENT_DIAB_T2) %>%
    ggplot(aes(x = value, color = INCIDENT_DIAB_T2)) +
    facet_wrap(~key) +
    geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=2, notch=FALSE) +
    xlim(-2, 2) +
    theme_classic() +
    theme(axis.title.x=element_blank(), axis.title.y=element_blank()) ) %>%
    ggsave(file = "cache/eicosanoid_boxplot.png",
           plot = .,
              height = 2,
              width = 10,
              dpi = 300)
```

<img src = "cache/eicosanoid_boxplot.png" />

</details>

<details><summary>Number of participants with missing values (NA)</summary>

```{r filtered values missing dilgom}
dset.dilgom.raw %>%
    mutate_all(is.na) %>%
    count_(vars = getmissing(mvars)) %>%
    kable
```

</details>

<details><summary>Number of participants with other exclusion criteria</summary>

```{r filtered values excluded dilgom}
dset.dilgom.raw %>%
    filter_at(vars(getmissing(mvars)), all_vars(!is.na(.))) %>%
    mutate_all(~. != 0) %>%
    count_(vars = getfilters(mvars)) %>%
    kable
```

</details>

### DILGOM characteristics

<details><summary>Continuous variables</summary>

```{r Characteristics dilgomcontinuous, echo = FALSE}
dset.dilgom %>%
    dplyr::select(female, one_of(getcontinuous(mvars))) %>%
    gather(key, value, -female) %>%
    group_by(key, female) %>%
    summarize(n = n(),
              na = sum(is.na(value)),
              mean = mean(value, na.rm = TRUE),
              sd = sd(value, na.rm = TRUE),
              gmean = gmean(value, na.rm = TRUE),
              iqr = IQR(value, na.rm = TRUE)) %>%
    mutate_if(is.numeric, ~round(., 3)) %>%
    kable
```

</details>

<details><summary>Categorical variables</summary>

```{r Characteristics cagegorical dilgom, echo = FALSE}
dset.dilgom %>%
    select(one_of(getcategorical(mvars) %difference% c("Sample_ID", "plate"))) %>%
    gather(key, value, -female) %>%
    group_by(key, female) %>%
    summarize(n = n(),
              na = sum(is.na(value)),
              zero = sum(value == 0, na.rm = TRUE),
              one = sum(value == 1, na.rm = TRUE),
              two = sum(value == 2, na.rm = TRUE),
              three = sum(value == 3, na.rm = TRUE)) %>%
    mutate_if(is.numeric, ~round(., 3)) %>%
    kable
```

</deails>


### Replication results in DILGOM 

<details>
  <summary>Running Cox models</summary>

```{r single eicosanoid models dilgom}
cox.dm.dilgom <-  loop.cox(dset = dset.dilgom %>% filter(PREVAL_DIAB_T2 == 0),
                           response = "DIAB_T2",
                           loops = getmetabolistes(dset.dilgom),
                           covariates = cox.covariates[["model_2"]])
```

```{r single eicosanoid models dilgom results}
cox.dm.dilgom  %>% loop.results(exponentiate = TRUE) %>% kable
```

</details>


### Risk score in DILGOM

<details>
  <summary>Risk score</summary>

```{r risk model riskclass dilgom, echo = FALSE}
rset.dilgom <- dset.dilgom %>%
    filter(PREVAL_DIAB_T2 == 0) %>%
    getriskset(., cox.fwd)
```

 ```{r risk model riskclass dilgom results, echo = FALSE}
lapply(c2l("riskclass", "riskpersd"), function(risk) {
    vars <- cox.covariates[["model_2"]] %difference% c("plate")
        loop.cox(dset = rset.dilgom,
                 response = "DIAB_T2",
                 loops = risk,
                 covariates = vars) }) %>%
    map_df(~tidy(.x, exponentiate = TRUE), .id = "model") %>%
    kable()
```

</details>

## Replication in FHS

```{r replication forest data}
forest.fhs <-
    tribble(~term, ~Identity, ~p.value, ~conf.low, ~conf.high, ~estimate,
            "mzid_279.196600_3.7247", "12-HHTrE", 0.066668527, 0.992064239, 1.27097149, 1.122891519,
            "mzid_311.223100_2.9230", "Unknown eicosanoid (311.2231/2.92)", 0.343607967, 0.802157423, 1.07983928, 0.93069925,
            "mzid_335.223200_2.6455", "8-iso-Prostaglandin A1", 0.000720178, 1.080055386, 1.335538077, 1.20102252) 
```

```{r meta-analysis fhs}
df.meta.fhs <-
    tribble(~term, ~estimate, ~conf.low, ~conf.high, ~std.error, ~p.value,
            "FHS", 1.236414, 1.101545, 1.387795, 0.05893022, 0.000317)
```

## Comparing replication results

```{r dilgom forest data, echo = FALSE}
forest.dilgom <- cox.dm.dilgom %>% loop.results(exponentiate = TRUE)
```

```{r fr02 forest data, echo = FALSE}
forest.fr02 <- cox.dm[["model_2"]]  %>%
    loop.results(exponentiate = TRUE) %>%
    filter(term %in% forest.fhs[["term"]])
```

```{r Comparing six-eicosanoid results}
dset.replication <- list(FINRISK = forest.fr02,
     FHS = forest.fhs,
     DILGOM = forest.dilgom) %>%
    map_df(identity, .id = "model") %>%
    arrange(term, match(model, c("FINRISK", "FHS", "DILGOM"))) %>%
    mutate(Identity = dplyr::recode(term, `mzid_279.196600_3.7247` = "12-HHTrE",
                               `mzid_311.223100_2.9230` = "Unknown eicosanoid (311.2231/2.92)",
                               `mzid_335.223200_2.6455` = "8-iso-Prostaglandin A1"),
           Identity = ifelse(model == "FINRISK", Identity, "")) %>% 
    mutate(mean_ci = sprintf("%.2f (%.2f to %.2f)", estimate, conf.low, conf.high))
```

```{r Plot the forest plot image, echo = FALSE
replicationforestplot(dset.replication, file = "cache/forestmap-replication.jpg")
```

<img src = "cache/forestmap-replication.png" />


## Meta-analysis

```{r meta-analysis fr02, echo = FALSE}
df.meta.fr02 <- loop.cox(dset = rset,
                         response = "DIAB_T2",
                         loops = "riskpersd",
                         covariates = cox.covariates[["model_2"]]) %>%
    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::filter(grepl("risk", term)) %>%
    select(-statistic) %>%
    mutate(term = "FINRISK")
```

```{r meta-analysis DILGOM, echo = FALSE}
df.meta.dilgom <- loop.cox(dset = rset.dilgom,
                         response = "DIAB_T2",
                         loops = "riskpersd",
                         covariates = cox.covariates[["model_2"]]) %>%
    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::filter(grepl("risk", term)) %>%
    select(-statistic) %>%
    mutate(term = "DILGOM")
```

```{r meta-analysis bind, echo = FALSE}
df.meta <- rbind(df.meta.fr02,
                 df.meta.dilgom,
                 df.meta.fhs)
```

Meta-analysis model

```{r meta-analysis model}
ret.meta <- metafor::rma(df.meta$estimate,
                         sei=df.meta$std.error,
                         method = "REML") %>%
    summary %>%
    coef %>%
    as.data.frame %>%
    mutate(term = "Overall")
```

Meta-analysis results

```{r meta-analysis results}
{ df.metaforest <- rbind(df.meta %>% select(term,
                                          estimate,
                                          conf.low,
                                          conf.high,
                                          p.value),
                         ret.meta %>% select(term,
                                             estimate,
                                             conf.low = ci.lb,
                                             conf.high = ci.ub,
                                           p.value = pval)) %>%
      mutate(mean_ci = sprintf("%.2f (%.2f to %.2f)", estimate, conf.low, conf.high),
             p.value = pub.p(p.value)) } %>%
    kable
```


```{r saving meta analysis plot, echo = FALSE}
jpeg(width = 4*1000, height = 4*420, res = 600, file = "cache/metaanalysis.png")
forestplot::forestplot(
                labeltext = cbind(c("Cohort", df.metaforest$term),
                                  c("HR (95% CI)", df.metaforest$mean_ci),
                                  c("P-value", df.metaforest$p.value)),
                mean = cbind(c(NA, df.metaforest$estimate)),
                lower = cbind(c(NA, df.metaforest$conf.low)),
                upper = cbind(c(NA, df.metaforest$conf.high)),
                is.summary = c(FALSE,rep(FALSE,nrow(df.metaforest)-1),TRUE),
                align = c("l", "l", "l"),
                graph.pos = 3,
                title = "",
                xlog = FALSE,
                xlab = "HR (95% CI)",
                hrzl_lines=list("2" = gpar(lwd=1, col="#000000")),
                txt_gp = fpTxtGp(label = gpar(cex = 1.5),
                                 summary = gpar(cex = 1.5, fontface = "plain"),
                                 ticks = gpar(cex = 1.5),
                                 xlab = gpar(cex = 1.5),
                                 title = gpar(cex = 1.5)),
                xticks = seq(0.0, 2.0, 0.5),
                clip =exp(c(-1, 1)),
                col = fpColors(box = "black", summary = "black"),
                zero = 1, 
                lineheight = unit(16, "mm"),
                boxsize = 0.2,
                colgap = unit(4, "mm"),
                lwd.ci = 1)
dev.off()
```

<img src = "cache/metaanalysis.png" />

