---
title: "Eicosanoids mediators of diabetes"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
---

```{r options, echo = FALSE}
options(knitr.kable.NA = "")
knitr::opts_chunk$set(include = TRUE,
                      echo = TRUE,
                      message = FALSE,
                      results = "asis",
                      cache = FALSE,
                      warning = FALSE)
```

# Libraries

<details>
  <summary>Open/Close</summary>

```{r libraries, results = 'hide'}
library(dplyr)
library(sessioninfo)
library(knitr)
library(readr)
library(tidyr)
library(parallel)
library(ggplot2)
library(broom)
library(gridExtra)
library(gplots)
library(survival)
library(purrr)
library(survminer)
library(tibble)
library(car)
library(forestplot)
library(rlang)
library(ggpubr)
library(reshape2)
library(metafor)
```

</details>

# Session

Analyses were performed using `r sessioninfo::platform_info()$version` on 
`r sessioninfo::platform_info()$system`.

<details><summary>Packages</summary>

```{r Session info, echo = FALSE}
sessioninfo::package_info() %>%
    data.frame %>%
    filter(attached == TRUE) %>%
    select(package, loadedversion, date, source) %>%
    kable(row.names = FALSE)
```

</details>

# Source code

```{r source file list, echo = FALSE}
sourcefiles <- c("articles-importer.R",
                 "articles-utils.R",
                 "articles-models.R",
                 "articles-plots.R",
                 "articles-step.R",
                 "articles-riskscore.R")
```

```{r import source files, echo = FALSE}
for (f in sourcefiles) {
    source(f)
}
```

```{r embed files to html, echo = FALSE}
xfun::embed_files(c("eicosanoidsdm.Rmd", sourcefiles))
```

#  Folder structure

```{r clean folders}
rmdir("cache")
mkdir("cache", "report", "session", "rds")
```

# Variables

<details><summary>Included variables</summary>

```{r variables to be include, echo = FALSE}
(mvars <- tribble(~var, ~type, ~dropmissing, ~useforfiltering,
                   "Sample_ID", "categorical", TRUE, FALSE,
                   "BL_AGE", "continuous", TRUE, FALSE,
                   "female", "binomial", TRUE, FALSE,
                   "PAINO", "continuous", TRUE, FALSE,
                   "PITUUS", "continuous", TRUE, FALSE,
                   "VYOTARO", "continuous", FALSE, FALSE,
                   "LANTIO", "continuous", FALSE, FALSE,
                   "BMI", "continuous", TRUE, FALSE,
                   "WHR", "continuous", FALSE, FALSE,
                   "SYSTM", "continuous", TRUE, FALSE,
                   "DIASM", "continuous", FALSE, FALSE,
                   "HBA1C", "continuous", FALSE, FALSE,
                   "ALKI2_FR02", "continuous", FALSE, FALSE,
                   "HDL", "continuous", FALSE, FALSE,
                   "LDL_DIRECT", "continuous", FALSE, FALSE,
                   "TRIG", "continuous", TRUE, FALSE,
                   "BP_TREAT", "binomial", TRUE, FALSE,
                   "KOULGR", "binomial", FALSE, FALSE,
                   "CURR_SMOKE", "binomial", FALSE, FALSE,
                   "Q57X", "categorical", TRUE, FALSE,
                   "hsCRP", "continuous", TRUE, FALSE,
                   "DIAB_FAMILYHIST", "binomial", TRUE, FALSE,
                   "PREVAL_DIAB_T2", "binomial", TRUE, FALSE,
                   "INCIDENT_DIAB_T2", "binomial", TRUE, FALSE,
                   "DIAB_T2_AGEDIFF", "continuous", TRUE, FALSE,
                   "EAST", "binomial", TRUE, FALSE,
                   "plate", "categorical", TRUE, FALSE,
                   "pregnant", "binomial", TRUE, TRUE,
                   "PREVAL_DIAB_T1", "binomial", TRUE, TRUE,
                   "INCIDENT_DIAB_T1", "binomial", TRUE, TRUE,
                   "INCIDENT_CR_ANYCANC", "binomial", TRUE, TRUE,
                  "PREVAL_CR_ANYCANC", "binomial", TRUE, TRUE,
                  "PREVAL_CVD", "binomial", FALSE, FALSE,
                  "LIPID_TREAT", "binomial", FALSE, FALSE) %>%
      dplyr::left_join(., importinfo(), by = "var")) %>%
    kable
```

</details>

# Data import

For the analyses, exclude pregnant women, prevalent/incident type 1 diabetics
and cancers (except non-melanoma skin cancers aka ICD10 category C44).

```{r import data}
dset <- (dset.raw <- importdata()) %>%
    filter_at(vars(getmissing(mvars)), all_vars(!is.na(.))) %>%
    filter_at(vars(getfilters(mvars)), all_vars(. == 0)) %>%
    mutate_at(vars(getcategorical(mvars)), factor)
```

```{r koe, eval = FALSE}
dset %>% pull(WHR) %>% summary
```

Dimensions before filtering are (`r dim(dset.raw)`) and after filtering (`r dim(dset)`).

```{r cox covariates, echo = FALSE}
( cox.covariates <- list("model_2" =  c("BL_AGE", "female", "EAST", "plate",
                                        "BMI", "Q57X", "DIAB_FAMILYHIST",
										"SYSTM", "BP_TREAT", "TRIG",
                                        "PREVAL_CVD", "LIPID_TREAT")) ) %>%
      map_df(~data.frame(covariates = paste(., collapse = ", ")), .id = "model") %>%
      kable
```

Running Cox models

```{r cox models}
cox.dm <- lapply(cox.covariates, function(covariates)
    loop.cox(dset = dset %>% filter(PREVAL_DIAB_T2 == 0),
             response = "DIAB_T2",
             loops = getmetabolistes(dset),
             covariates = covariates))
```

```{r correlation model last element, echo = FALSE}
coxmodel.significant.eicosanoids <- cox.dm[["model_2"]] %>%
    loop.results %>%
    filter(qval < 0.05)
```

# Stepwise Cox regression


```{r cox forward dataset}
dset.fwd <- dset %>%
    filter(PREVAL_DIAB_T2 == 0) %>%
#    mutate_at(vars(getcategorical(mvars) %union% "plate" %difference% "INCIDENT_DIAB_T2"), as.numeric) %>%
    mutate_at(vars("INCIDENT_DIAB_T2"), eventasnumber)
cox.fwd.mzids <- coxmodel.significant.eicosanoids %>% pull(term)
```

```{r stepwise covariates}
( risk.covariates <- list(risk1 = cox.covariates[["model_2"]],
                          risk2 = cox.covariates[["model_2"]] %difference% "BMI" %union% "WHR",
                          risk3 = cox.covariates[["model_2"]] %difference% "BMI" %union% "VYOTARO")) %>% 
    map_df(~data.frame(covariates = paste(., collapse = ", ")), .id = "model") %>%
    kable
```

```{r cox normal, results = 'hide'}
cox.dm <- lapply(risk.covariates, function(covariates)
    loop.cox(dset = dset %>% filter(PREVAL_DIAB_T2 == 0),
             response = "DIAB_T2",
             loops = c("mzid_279.196600_3.7247", "mzid_311.223100_2.9230", "mzid_335.223200_2.6455"),
             covariates = covariates))
```

```{r cox normal results, echo = FALSE}
cox.dm %>%
    map_df(~loop.results(.x), .id = "model") %>%
    filter(qval < 0.05) %>%
    mutate(result = sprintf("%.3f ± %.3f (P=%.3e)", estimate, std.error, p.value)) %>%
    mutate(term = dplyr::recode(term, `mzid_279.196600_3.7247` = "12-HHTrE",
                                    `mzid_311.223100_2.9230` = "Unknown eicosanoid (311.2231/2.92)",
                                    `mzid_335.223200_2.6455` = "8-iso-Prostaglandin A1")) %>% 
    select(model, term, result) %>%
    spread(model, result) %>%
    kable
```


```{r cox forward, results = 'hide'}
cox.fwd <- lapply(risk.covariates, function(covariates)
    stepwise.coxph(Time = "DIAB_T2_AGEDIFF",
                   Status = "INCIDENT_DIAB_T2",
                   variable.list = cox.fwd.mzids,
                   in.variable = covariates,
                   sle = 0.05/length(getmetabolistes(dset)),
                   sls = 0.05/length(getmetabolistes(dset)),
                   data = dset.fwd))
```

```{r koe, eval = FALSE, include = FALSE}

dset.fwd %>% pull(WHR) %>% summary

coxph(formula = Surv(DIAB_T2_AGEDIFF, INCIDENT_DIAB_T2) ~  
          koe,  
    data = dset.fwd %>% mutate(koe = LANTIO/VYOTARO))

```

<details><summary>Numerical results</summary>

```{r cox fwd results, echo = FALSE}
cox.fwd %>%
    map_df(~tidy(.x, exponentiate = TRUE), .id = "model") %>%
    kable
```

</details>

```{r cox check proportionality}
cox.fwd[["risk2"]]
cox.fwd[["risk2"]]
%>%
    check.proportionality

    map_df(~as.data.frame(.x$table) %>% tibble::rownames_to_column("term"),
           .id = "testing") %>%
    kable
```

## Risk score

```{r risk score}
df.risk_score <- lapply(c2l(names(risk.covariates)), function(id) {
    rset <- dset %>%
        filter(PREVAL_DIAB_T2 == 0) %>%
        getriskset(., cox.fwd[[id]])
    loop.cox(dset = rset,
             response = "DIAB_T2",
             loops = "riskpersd",
             covariates = risk.covariates[[id]])
})
```


```{r risk model riskclass, echo = FALSE}
df.risk_score %>%
    map_df(~tidy(.x, exponentiate = TRUE), .id = "model") %>%
    filter(term == "riskpersd") %>%
    mutate(p.value = pub.p(p.value)) %>% 
    kable
```

# No preval CVD

```{r cox forward no preval cvd, results = 'hide'}
cox.fwd_nopreval <- stepwise.coxph(Time = "DIAB_T2_AGEDIFF",
                   Status = "INCIDENT_DIAB_T2",
                   variable.list = cox.fwd.mzids,
                   in.variable = cox.covariates[["model_2"]] %difference% "PREVAL_CVD",
                   sle = 0.05/length(getmetabolistes(dset)),
                   sls = 0.05/length(getmetabolistes(dset)),
                   data = dset.fwd %>% filter(PREVAL_CVD == 1))
```

<details><summary>Numerical results</summary>

```{r cox fwd results no preval cvd, echo = FALSE}
cox.fwd_nopreval %>%
    tidy(exponentiate = TRUE) %>% 
    kable
```

</details>

## Risk score

```{r risk score no preval cvd}
rset <- dset %>%
    filter(PREVAL_DIAB_T2 == 0) %>%
    getriskset(., cox.fwd_nopreval)

loop.cox(dset = rset,
         response = "DIAB_T2",
         loops = "riskpersd",
         covariates = cox.covariates[["model_2"]] %difference% "PREVAL_CVD") %>%
    tidy(exponentiate = TRUE) %>% 
    filter(term == "riskpersd") %>%
    mutate(p.value = pub.p(p.value)) %>% 
    kable
```



